---
title: "R Notebook"
output: html_notebook
---

This code imports, analyzes, and plots data from a multiple replcicates of beta-galactosidase assays. This analysis is for experiments examining beta-galactosidase activity as a function of RpoS level.

Before running this code, you must have followed the directions and run the code in "analysis of single beta-gal experiment.Rmd" on every file you intend analyze here. 

```{r, message=FALSE}
library(readxl)
library(tidyverse)
source("betaGalAnalysisFunctions.R")
theme_set(theme_classic())


#Data taken on 7-18-18 and 7-19-18 are preliminary and should not be included in final analysis. Data taken on 7-25-18 and 7-26-18 are from Liya's frozen strains. All other data are from Francisco's frozen strains. I've created different variations of 'fileNames' to run code on subsets of the data. 

#gadB wt (CRD38) and astC wt (CRD39)
#fileNames <- c("CRDdata/CRD 7-17-18 OD420 and OD600 afternoon.xlsx", "CRDdata/CRD 7-18-18 OD420 and OD600 morning.xlsx", "CRDdata/CRD 7-19-18 OD420 and OD600 afternoon.xlsx", "CRDdata/RAN 7-24-18 BETA GAL RAN11, RAN12, RAN13.xlsx", "CRDdata/RAN 7-26-18 BETA GAL RAN11, RAN12, RAN13.xlsx", "CRDdata/RAN 7-27-18 BETA GAL RAN11, RAN12.xlsx", "CRDdata/RAN 7-28-18 BETA GAL RAN11, RAN12, RAN13.xlsx", "CRDdata/RAN 7-29-18 BETA GAL RAN11, RAN12, RAN13.xlsx")

#astC controls
#fileNames <- c("myData/LEB 7-20-18 Strain 41 42 43.xlsx", "myData/LEB 7-25-18 Strain 41 42 43.xlsx", "myData/LEB 7-26-18 Strain 41 42 43.xlsx", "myData/LEB 7-27-18 Strain 41 42 43.xlsx", "myData/LEB 7-30-18 Strain 41 42 43.xlsx", "myData/LEB 7-31-18 Strain 41 42 43.xlsx")
   
#gad controls         
#fileNames <- c("FRdata/7.19.18 FRLZ Strain 37 38.xlsx", "FRdata/7.19.18 FRLZ Strain 39 40.xlsx", "FRdata/7.20.18 FRLZ Strain 37 38.xlsx", "FRdata/7.20.18 FRLZ Strain 39 40.xlsx", "FRdata/7.25.18 FRLZ Strain 37 38.xlsx", "FRdata/7.25.18 FRLZ Strain 39 40.xlsx", "FRdata/7.26.18 FRLZ Strain 37 38.xlsx", "FRdata/7.26.18 FRLZ Strain 39 40.xlsx", "FRdata/7.27.18 FRLZ Strain 37 38.xlsx", "FRdata/7.27.18 FRLZ Strain 39 40.xlsx", "CRDdata/CRD 8-1-18 Strain 37 38 39 40.xlsx")

#gad and ast controls
#fileNames <- c("FRdata/7.19.18 FRLZ Strain 37 38.xlsx", "FRdata/7.19.18 FRLZ Strain 39 40.xlsx", "FRdata/7.20.18 FRLZ Strain 37 38.xlsx", "FRdata/7.20.18 FRLZ Strain 39 40.xlsx", "FRdata/7.25.18 FRLZ Strain 37 38.xlsx", "FRdata/7.25.18 FRLZ Strain 39 40.xlsx", "FRdata/7.26.18 FRLZ Strain 37 38.xlsx", "FRdata/7.26.18 FRLZ Strain 39 40.xlsx", "FRdata/7.27.18 FRLZ Strain 37 38.xlsx", "FRdata/7.27.18 FRLZ Strain 39 40.xlsx", "myData/LEB 7-20-18 Strain 41 42 43.xlsx", "myData/LEB 7-25-18 Strain 41 42 43.xlsx", "myData/LEB 7-26-18 Strain 41 42 43.xlsx", "myData/LEB 7-27-18 Strain 41 42 43.xlsx", "myData/LEB 7-30-18 Strain 41 42 43.xlsx", "myData/LEB 7-31-18 Strain 41 42 43.xlsx", "CRDdata/CRD 8-1-18 Strain 37 38 39 40.xlsx")

#astC mutants
#fileNames <- c("myData/LEB 7-20-18 Strain 34 35 36.xlsx", "myData/LEB 7-25-18 Strain 34 35 36.xlsx", "myData/LEB 7-26-18 Strain 34 35 36.xlsx", "myData/LEB 7-27-18 Strain 34 35 36.xlsx", "myData/LEB 7-30-18 Strain 34 35 36.xlsx", "myData/LEB 7-31-18 Strain 34 35 36.xlsx")

#gadB mutants
#fileNames <- c("FRdata/7.19.18 FRLZ Strain 31 33.xlsx","FRdata/7.19.18 FRLZ Strain 32 30.xlsx", "FRdata/7.20.18 FRLZ Strain 31 33.xlsx", "FRdata/7.20.18 FRLZ Strain 32 30.xlsx", "FRdata/7.23.18 FRLZ Strain 31 33.xlsx", "FRdata/7.23.18 FRLZ Strain 32 30.xlsx", "FRdata/7.26.18 FRLZ Strain 31 33.xlsx", "FRdata/7.26.18 FRLZ Strain 32 30.xlsx", "FRdata/7.27.18 FRLZ Strain 31 33.xlsx", "FRdata/7.27.18 FRLZ Strain 32 30.xlsx", "myData/LEB 8-1-18 Strain 30 31 32 33.xlsx")

#gadB mutants + wt
fileNames <- c("FRdata/7.19.18 FRLZ Strain 31 33.xlsx","FRdata/7.19.18 FRLZ Strain 32 30.xlsx", "FRdata/7.20.18 FRLZ Strain 31 33.xlsx", "FRdata/7.20.18 FRLZ Strain 32 30.xlsx", "FRdata/7.23.18 FRLZ Strain 31 33.xlsx", "FRdata/7.23.18 FRLZ Strain 32 30.xlsx", "FRdata/7.26.18 FRLZ Strain 31 33.xlsx", "FRdata/7.26.18 FRLZ Strain 32 30.xlsx", "FRdata/7.27.18 FRLZ Strain 31 33.xlsx", "FRdata/7.27.18 FRLZ Strain 32 30.xlsx", "myData/LEB 8-1-18 Strain 30 31 32 33.xlsx", "CRDdata/CRD 7-17-18 OD420 and OD600 afternoon.xlsx", "CRDdata/CRD 7-18-18 OD420 and OD600 morning.xlsx", "CRDdata/CRD 7-19-18 OD420 and OD600 afternoon.xlsx", "CRDdata/RAN 7-24-18 BETA GAL RAN11, RAN12, RAN13.xlsx", "CRDdata/RAN 7-26-18 BETA GAL RAN11, RAN12, RAN13.xlsx", "CRDdata/RAN 7-27-18 BETA GAL RAN11, RAN12.xlsx", "CRDdata/RAN 7-28-18 BETA GAL RAN11, RAN12, RAN13.xlsx", "CRDdata/RAN 7-29-18 BETA GAL RAN11, RAN12, RAN13.xlsx")

#astC mutants + wt
#fileNames <- c("myData/LEB 7-20-18 Strain 34 35 36.xlsx", "myData/LEB 7-25-18 Strain 34 35 36.xlsx", "myData/LEB 7-26-18 Strain 34 35 36.xlsx", "myData/LEB 7-27-18 Strain 34 35 36.xlsx", "myData/LEB 7-30-18 Strain 34 35 36.xlsx", "myData/LEB 7-31-18 Strain 34 35 36.xlsx", "CRDdata/CRD 7-17-18 OD420 and OD600 afternoon.xlsx", "CRDdata/CRD 7-18-18 OD420 and OD600 morning.xlsx", "CRDdata/CRD 7-19-18 OD420 and OD600 afternoon.xlsx", "CRDdata/RAN 7-24-18 BETA GAL RAN11, RAN12, RAN13.xlsx", "CRDdata/RAN 7-26-18 BETA GAL RAN11, RAN12, RAN13.xlsx", "CRDdata/RAN 7-27-18 BETA GAL RAN11, RAN12.xlsx", "CRDdata/RAN 7-28-18 BETA GAL RAN11, RAN12, RAN13.xlsx", "CRDdata/RAN 7-29-18 BETA GAL RAN11, RAN12, RAN13.xlsx")

processedFiles <- map(fileNames, meanExpressionOneReplicate)

combinedData <- bind_rows(processedFiles)
```

Plot lacZ expression vs. RpoS for each strain
```{r}

ggplot(combinedData) + 
  geom_point(aes(x=RpoS, y=meanExpression)) + 
  facet_wrap(~strain, scales = "free") +
  labs(x = "% of wild-type RpoS", y = "LacZ expression (Miller Units)") + 
  expand_limits(x=0, y=0) 

```

Plot mean expression ± SE for each RpoS level as well as an overlaid plot of expression profiles for all strains.
```{r}
dayMeans <- combinedData %>% 
              group_by(strain, RpoS) %>% 
              #filter(strain =="FR30") %>% #use this line to view individual strains
              summarise(meanAcrossDays = mean(meanExpression), sdAcrossDays = sd(meanExpression)/sqrt(length(meanExpression))) %>%
              mutate(upperErrorBar = meanAcrossDays + sdAcrossDays, lowerErrorBar = meanAcrossDays - sdAcrossDays)

ggplot(dayMeans) + 
  geom_point(aes(x = RpoS, y = meanAcrossDays)) + 
  geom_errorbar(aes(x = RpoS, y = meanAcrossDays, ymax = upperErrorBar, ymin = lowerErrorBar, width = 8)) + 
  facet_grid(.~strain, scales = "free") +
  labs(x = "% of wild-type RpoS", y = "LacZ expression (Miller Units)")

#this plots the mean value for all strains, sorted by color 
ggplot(dayMeans) + 
  geom_point(aes(x=RpoS, y=meanAcrossDays, color=strain)) +
  labs(x = "% of wild-type RpoS", y = "LacZ expression (Miller Units)")

  
```


##Sensitivity

This section is specific for analyzing the sensitivity of a promoter. It requires three RpoS levels.

This function expects three RpoS levels, so for this example I'll subset out some, since we have five.

```{r}
dataForSensitivity <- combinedData %>% 
                        filter(RpoS == 0 | RpoS == 26 | RpoS == 89) 

```


Calculate sensitivity of a promoter. 
```{r}
sensitivities <- dataForSensitivity %>% 
                  group_by(strain, date, experimenter) %>%
                  summarise(sensitivity = calculateSingleSens(RpoS, meanExpression))

```


Plot the data with a line to illustrate null expectation used to assess sensitivity.

```{r}
segmentEnds <- dayMeans %>%
                filter(RpoS == 0 | RpoS == 89) %>%
                group_by(strain) %>%
                summarise(
                  RpoS_min = min(RpoS),
                  RpoS_max = max(RpoS),
                  expression_min = min(meanAcrossDays),
                  expression_max = max(meanAcrossDays)
                  )
```


If you designed your experiment with only three RpoS levels, delete the the `filter` step. Note that we have to go back a step to use use the `dayMeans` tibble.
```{r}


#rename strains with genotype and store in 'genotype' variable
genotype <- as_labeller(c(`LZ34` = "ΔargR", `LZ35` = "Δnac", `LZ36` = "ΔntrC", `LZ41` = "argR control", `LZ42` = "nac control", `LZ43` = "ntrC control", `FR30` = "ΔgadX", `FR31` = "ΔgadW", `FR32` = "ΔgadXYW", `FR33` = "ΔgadE", `FR37` = "gadX control", `FR38` = "gadW control", `FR39` = "gadXYW control", `FR40` = "gadE control", `CRD38` = "gadB wt", `CRD39` = "astC wt"))

#plot mean expression ± SE with a line
dayMeans %>% 
  ggplot() + 
  geom_point(aes(x = RpoS, y = meanAcrossDays), show.legend = FALSE) + 
  
  #creates error bars 
  geom_errorbar(aes(x = RpoS, y = meanAcrossDays, ymax = upperErrorBar, ymin = lowerErrorBar, width  = 8)) + 
  
  #facet_wrap creates a grid that allows us to define dimensions
  facet_wrap(~strain, scales = "free", ncol = 2, labeller = genotype) +
  
  #this formats the plot titles to remove border and position text outside plot
  theme(aspect.ratio = 0.5) +
  xlim(0, 100) +
  theme(strip.background = element_blank(), strip.placement = "outside") +
  
  #this function plots the line to illustrate null expectation
  geom_segment(data = segmentEnds, mapping = aes(x = RpoS_min, xend = RpoS_max, y = expression_min, yend = expression_max)) +
  
  #add correct axis titles and coordinates
  labs(x = "% of wild-type RpoS", y = "LacZ expression (Miller Units)") +
  
  #axis begins at (0,0)
  expand_limits(x = 0, y = 0) 

```


This chunk plots lacZ expression vs. mutation facetted by RpoS level. You'll need to manually run this chunk. 

```{r, eval = FALSE}
ggplot(combinedData) + 
  geom_point(aes(x=strain, y=meanExpression)) + 
  facet_grid(~RpoS) +
  labs(x = "Mutation", y = "LacZ expression (Miller Units)") + 
  expand_limits(x=0, y=0) + 
  theme(axis.text.x=element_text(angle = -90, hjust = 0)) +
  scale_x_discrete(labels = c("gadX","gadW","gadxYW","gadE", "argR", "nac", "ntrC"))
```

