---
title: "Untitled"
author: "Lakshmi Batachari"
date: "July 19, 2015"
output: word_document
---
```{r, message = FALSE}
library(readxl)
library(tidyverse)
theme_set(theme_classic())
```

Import your files here:
```{r, message = FALSE}

#gadB mutants
fileNames <- c("FRdata/7.19.18 FRLZ Strain 31 33.xlsx","FRdata/7.19.18 FRLZ Strain 32 30.xlsx", "FRdata/7.20.18 FRLZ Strain 31 33.xlsx", "FRdata/7.20.18 FRLZ Strain 32 30.xlsx", "FRdata/7.23.18 FRLZ Strain 31 33.xlsx", "FRdata/7.23.18 FRLZ Strain 32 30.xlsx", "FRdata/7.26.18 FRLZ Strain 31 33.xlsx", "FRdata/7.26.18 FRLZ Strain 32 30.xlsx", "FRdata/7.27.18 FRLZ Strain 31 33.xlsx", "FRdata/7.27.18 FRLZ Strain 32 30.xlsx")

processedFiles <- map(fileNames, meanExpressionOneReplicate)

combinedData <- bind_rows(processedFiles)
```

Create vectors/variables we'll need in the next step. 
```{r}
#set number of randomizations
sampleSize <- 10000

#Set number of biological replicates
repSize <- 5

#create vector for the 26% RpoS level estimates
linear26Est <- vector(length=sampleSize)

#define strain
Strain <- "FR33"

#store mean expression values for the selected strain at 0%, 26%, and 89%
rpoS0 <- subset(combinedData, RpoS == 0 & strain == Strain, select = c(meanExpression))
rpoS26 <-subset(combinedData, RpoS == 26 & strain == Strain, select = c(meanExpression))
rpoS89 <-subset(combinedData, RpoS== 89 & strain == Strain, select = c(meanExpression))

#convert expression values into vector
rpos0Vect <- as.numeric(as.character(rpoS0$meanExpression))
rpos26Vect <- as.numeric(as.character(rpoS26$meanExpression))
rpos89Vect <- as.numeric(as.character(rpoS89$meanExpression))

#create vectors to store randomly sampled expression values at 0% and 89%
zeroPercentC <- vector(mode = "numeric", length= repSize)
eightyNinePercentC <- vector(mode = "numeric", length= repSize)
```


Now we sample from the data 10,000 times:
```{r}
#conduct resampling 10,000 times
for (i in 1:sampleSize) 
{
replicates <- sample(c(1:repSize), replace = TRUE, repSize) #sample numbers the same number of times as biological replicates with replacement. This vector will be used to randomly pick an expression value from the vector containing all our of experimental values

#randomly samples biological replicate number expression values each at 0% and 89%
for (j in 1:repSize) 
{
#choose an expresion value from 0% and 89% RpoS by indexing with the number that was randomly sampled and stored in the vector 'replicates'
  
zeroPercentC[j] <- rpos0Vect[replicates[j]] 
eightyNinePercentC[j] <- rpos89Vect[replicates[j]] 

}
#determines median of expression values at the 0% and 89% level
zeroPercent <- median(zeroPercentC) 
eightyNinePercent <- median(eightyNinePercentC)

#obtains a linear fit from the median of the sampled 0% level to the median of the 89% level 
fit <- lm(c(zeroPercent, eightyNinePercent) ~ c(0, 89))

#estimates expression at the 26% RpoS level using the slope and y-intercept of the linear fit, stores estimates in vector
estimate <- 26*coef(fit)[2] + coef(fit)[1]
linear26Est[i] <- estimate
}
```


Now we have a distribution of expression values at 26% as predicted by our linear fit. We can use this distribution to determine the upper and lower bounds of the 95% confidence interval and the p-value for sensitivity. The null hypothesis is that the expression profile is linear. 

```{r}
#sorts 26% level estimates from lowest to highest
linear26Est <- sort(linear26Est)

#calculates lower and upper bounds for the 95% confidence interval of the expression at 26% RpoS and prints
lowerBound <- linear26Est[sampleSize*0.025]
upperBound <- linear26Est[sampleSize*0.975]

lowerBound
upperBound

#calculates p-value for sensitivity by determining probability that the median 26% RpoS value will fall outside of the 95% confidence interval
if (median(rpos26Vect) > median(linear26Est)){ #checks so that a p-value will always be generated instead of 1-p
  1-sum(median(rpos26Vect) > linear26Est)/sampleSize
} else {
  sum(median(rpos26Vect) > linear26Est)/sampleSize
}

#show histogram expression at 26% as predicted by linear fit
hist(linear26Est)


```

